Contention Management for 5G

### 2017-02-16 article/article.tex
relecture + updates

      A {\bf $(P,\tau)$-periodic affectation} of a routage graph $\cal R$ is a sequence  ${\cal M}=(m_0, \ldots ,m_{c-1})$ of $c$ integers that we call {\bf offsets}, with $c$ the number of routes in $\cal R$. The number $m_i$ represents the index of the first slot used in a period  by the route $r_i \in {\cal R}$ at its source.
      A $P$-periodic affectation must have no {\bf collision} between two routes in ${\cal R}$, that is $\forall (r_i, r_j) \in {\cal R}^2, i \neq j$, % with $\tau$ the size (in number of consecutive slots) of each message that must be periodically sent on each route of ${\cal R}_{\cal C}$, 
      we have $$[t(u,r_i)] \cap [t(u,r_j)] = \emptyset .$$

%       Notice that the notion of $P$-periodic affectation \textbf{is not monotone} with regard to $P$. 
      As an exemple of a $(2,1)$-periodic affectation, let consider a routage graph with routes $\{r_i\}_{i=1,\dots,c}$, such that all pairs of routes intersect at a different edge.
      We set $\tau = 1$ and the delays are chosen so that if $r_i$ and $r_j$ have $v$ as first common vertex then we have $\lambda(v,r_i) - \lambda(v,r_j)=1$.There is a $(2,1)$-periodic affectation by setting all $m_i$ to $0$.

@@

      \noindent {\bf Input:} a routage graph $\cal R$, an integer $\tau$ and an integer $P$.

      \noindent {\bf Question:} does there exist a $(P,\tau)$-periodic affectation of $\cal R$ ?

@@

      Consider the routage graph ${\cal R}$ given in the previous subsection. 
      We change the delays so that for $v$, the first vertex which belongs to $r_i$ and $r_j$,
      we have $\lambda(v,r_i) - \lambda(v,r_j)= P$, where $P$ is an odd number smaller than $c$ the number of routes in ${\cal R}$.
      If we consider a period of $2$, for all $i \neq j$, $\lambda(v,r_i) - \lambda(v,r_j) = 1 $ modulo $2$. Therefore $(0,\dots,0)$ is a $(2,1)$-periodic affectation of ${\cal R}$.

@@

      In the context of cloud-RAN applications, we consider here the digraph $G=(V,A)$ modeling the target network 
      and two disjoint subsets of vertices $S$ and $T$ of equal cardinality, where $S$ is the set of RRHs and $T$ is the set of BBUs. 
      A \textbf{symmetric} assignment ${\cal C}$, is an involutive function from $S$ to $T$, which maps each element $s\in S$ to ${\cal C}(s)$ and ${\cal C}(s)$ to $s$. It can also be seen as the set of pairs $(s,{\cal C}(s))$ and $({\cal C}(s),s)$.

       We are given a period $P$, a routing function ${\cal R}$, and we consider a $(P,\tau)$-periodic affectation of ${\cal R}_{\cal C}$ which associates $m_s$ to the route $r_s$ which begins by $s$.  This affectation represents the following process: first a message is sent at $s$, through the route $r_s$, at time $m_s$.

@@

      \begin{center}
      \includegraphics[scale=0.2]{rrh.jpeg}
      \end{center}

@@

       \begin{center}

      \includegraphics[scale=0.2]{BBU2.jpeg}
      \end{center}

@@

\section{Solving PRA}
  \label{sec:complexity}
  \subsection{NP-Hardness}

 In this section we assume that the size of a message $\tau$ is equal to one. 
 We will prove the hardness of PRA and PALL for this parameter, which implies the hardness of the general problems. 
Consider an instance of the problem PRA, i.e., a routage graph $\cal{R}$, a message size $\tau$ and a period $P$. 
The {\bf conflict depth} of a route is the number of other routes which share an edge with it. 
The conflict depth of a routage graph  $\cal{R}$ is the maximum of the conflict depth of the routes in ${\cal R}$.
The {\bf load} of a routage graph is the maximal number of routes sharing the same arc.
It is clear that a $(P,\tau)$-periodic affectation must satisfy that $P$ is larger or equal to the load times $\tau$.

@@

      \Vertex[x=12,y=3]{l3}
      \Vertex[x=14,y=4]{l2}
      \Vertex[x=10,y=2]{l1}
      \tikzstyle{VertexStyle}=[shape = circle, draw, minimum size = 20pt]
	\tikzset{
      VertexStyle/.append style = {blue} }
	\Vertex[x=-8,y=3]{1}
	      \tikzset{
      VertexStyle/.append style = {green} }
	  \Vertex[x=-7,y=5]{2}

@@

      \SetVertexNoLabel
      \Vertex[x=2,y=5]{A}
      \Vertex[x=4,y=5]{B}
      \Vertex[x=10,y=5]{C}
      \Vertex[x=12,y=5]{D}
      \Vertex[x=6,y=3]{E}
      \Vertex[x=8,y=3]{F}
      \tikzset{
      EdgeStyle/.append style = {green} }
      \Edge[label = 3](s2)(A)
      \Edge(A)(B)
      \Edge(B)(C)
      \Edge(C)(D)
      \Edge(D)(l2)

      \tikzset{
      EdgeStyle/.append style = {red} }
      \Edge[label = 1](s3)(E)
      \Edge(E)(F)
      \Edge(F)(l3) 
	\tikzset{
      EdgeStyle/.append style = {blue} }
      \Edge[label = 1](s1)(A)
      \Edge[label = 1](A)(B)
      \Edge[label = 1](B)(E)
      \Edge(E)(F)
      \Edge(F)(l1)

@@

      In this section, we consider graphs with a very simple topology that we call the {\bf star topology}. 
      First, for each arc $(u,v)$, there is also an arc $(v,u)$ with the same weight.
      Moreover, there is a special arc, the central arc, which is shared by all routes.
      All routes consist from an arc from its source to the source node, denoted by {\bf SN},
      then the central arc to {\bf TN}, the target node, and an arc to its target. In addition to SN and TN,
      there are two sets $S$ and $T$ of cardinality $n$ and ${\cal C}$ a symmetric assignment from $S$ to $T$. 
      The routes are the directed paths of the form $s_i,SN,TN,t_i$ and $t_i,TN,SN,s_i$.

@@

      We assume the weight of the central arc to be $0$ since it appears in every route,
      its value does not matter when considering affectations. Collisions between messages can 
      only appear at node SN on the way forward and at node TN on the way back. 
  \todo{Définir la forward period et la backward period et dire à quoi correspond une affectation en terme de forward et backward period (placement de messages dans la première période}     
%       Therefore, 
%       finding an $(P,\tau)$-periodic affectation without collision, is equivalent to fix ...

@@

  In this subsection, we deal with a simpler version of the problem PALL.
  We ask for a $(P,\tau)$-periodic affectation with all waiting times equal to $0$. 
  In that case $T_{max}$ is not relevant anymore. Since $w_i=0$, choosing $m_i$, the offset of the route from
  $s_i$ to $t_i$, also sets the offset of the route from $t_i$ to $s_i$ to $m_i + t(r_i)$.
  Moreover, in this context the weight of the arcs from $s_i$ to SN have no impact, therefore we assume they are 
  equal to $0$.

@@

    \begin{proposition}
    There is an algorithm, which finds a $(P,\tau)$-periodic affectation in time $O(n^2)$ when $P \geq 3n\tau$.
    \end{proposition}
    \begin{proof}
     We consider the time period at the vertex SN and cut it into consecutive macro slots of time of size $\tau$. The algorithm works in the following way: select the first route for which the offset as not been chosen. Try every offset which put the message in a  
     free macro slot in the period at SN. It also fixes the position of a message a TN, and we choose the first offset which does not create a collision at TN. 
     Assume we choose the offset of the route $r_{k+1}$, we have  at least $3n - k$ free macro-slots, since $P \geq 3n\tau$. Each of these $3n - k$ possible offset values translates into $3n - k$ positions of messages in the period at TN. All these positions are separated by at least $\tau$ slots. There are already $k$ messages of size $\tau$ in the period at TN. One such message can intersect at most $2k$ potential positions, therefore  amongst the possible $3n - k$ positions, there are  $3n - k -2k$ which are without collision. Since $k < n$, one of the choice is without collision, which proves that the algorithm terminates and find a
     $(P,\tau)$-periodic affectation. To ensure a complexity linear in $n$, it is enough to maintain a list of the free slices at SN and a list of the free intervals at TN. 
    \end{proof}

@@

	\begin{algorithm}[H]
	\caption{Greedy affectation}
	\begin{algorithmic}
	\REQUIRE ${\cal R}_{\cal C}$, period $P$
	\ENSURE A P-periodic affectation in p $\leq P$, or FAILURE
	\STATE $P1[P]$ slots of size $\tau$ in first way period.
	\STATE $P2[P]$ slots  of size $\tau$ in back way period.
	\FORALL{route i in ${\cal R}_{\cal C}$ }

@@

	\end{algorithmic}
	\end{algorithm}
      \paragraph{Period}
	This algorithm gives us a solution without waiting times in a maximum period $3.\tau.c$, if we have $c$ routes.\\
	Suppose that we have a period of $3.\tau.c$ slots, divided in $\tau$ macro-slots. Put a message in the first slot of size $\tau$ in the forward period, such that the corresponding area in the backward period is free.This message takes at most 2 slots of time $\tau$ in the backward period.\\
      \begin{center}
      \includegraphics[scale=0.2]{ex3nt.png}
      \end{center}
	When $k < c$ messages are put in the forward period, and we want to add another message, there is $3.c - k$ free slots of size $\tau$ in the forward period. Those $3.c - k$ gives us $3.c - k$ possible slots in the backward periods.\\
	The $k$ messages uses at most $2k$ slots of size $\tau$ used in the backward period. Since $k<c$,  $2k < 3.c - k$, thus using the pigeonhole principle, there is at least one free slot in the backward period for the new message.

    \subsubsection{Exhaustive generation}
	\begin{algorithm}[H]
	\caption{Exhaustive Generation}  
	\begin{algorithmic}
	\REQUIRE A routage graph ${\cal R}_{\cal C}$, period $P$, packet size $\tau$
	\ENSURE $(P,\tau)$-periodic affectation of ${\cal R}_{\cal C}$
	\STATE Forward-budget $\leftarrow$ $P$ - {\cal c} * $\tau$
	\STATE Backward-budget $\leftarrow$ $P$ - {\cal c} * $\tau$
	\STATE Free-Intervals $\leftarrow$ list of free intervals, init to $[0;P[$
	\FORALL{route i in  ${\cal R}_{\cal C}$}
	\FORALL{j in Free-Intervals }
	\IF{Message of the route i does not collides with scheduled routes}
	\STATE $m_i \leftarrow $ the first slot of Free-Intervals[j]
	\STATE Split the Free-Intervals considering the new packet
	\STATE Forward-budget $\leftarrow$ Forward-budget - {\em lost size}
	\STATE Backward-budget $\leftarrow$ Backward-budget - {\em lost size}
	\STATE call Exhaustive Generation on remaining routes
	\ENDIF
	\ENDFOR
	\ENDFOR

>>>>

      A {\bf $(P,\tau)$-periodic affectation} of a routage graph $\cal R$ is a sequence  ${\cal M}=(m_0, \ldots ,m_{c-1})$ of $n$ integers that we call {\bf offsets}, with $n$ the number of routes in $\cal R$. The number $m_i$ represents the index of the first slot used in a period  by the route $r_i \in {\cal R}$ at its source.
      A $(P,\tau)$-periodic affectation must have no {\bf collision} between two routes in ${\cal R}$, that is $\forall (r_i, r_j) \in {\cal R_{\cal C}}^2, i \neq j$, % with $\tau$ the size (in number of consecutive slots) of each message that must be periodically sent on each route of ${\cal R}_{\cal C}$, 
      we have $$[t(u,r_i)] \cap [t(u,r_j)] = \emptyset .$$

%       Notice that the notion of $P$-periodic affectation \textbf{is not monotone} with regard to $P$. 
      As an example of a $(2,1)$-periodic affectation, let consider a routage graph with routes $\{r_i\}_{i=1,\dots,n}$, such that all pairs of routes intersect at a different edge.
      We set $\tau = 1$ and the delays are chosen so that if $r_i$ and $r_j$ have $v$ as first common vertex then we have $\lambda(v,r_i) - \lambda(v,r_j)=1$.There is a $(2,1)$-periodic affectation by setting all $m_i$ to $0$.

@@

      \noindent {\bf Input:} a routage graph $\cal R_{\cal C}$, an integer $\tau$ and an integer $P$.

      \noindent {\bf Question:} does there exist a $(P,\tau)$-periodic affectation of $\cal R_{\cal C}$ ?

@@

      Consider the routage graph ${\cal R_{\cal C}}$ given in the previous subsection. 
      We change the delays so that for $v$, the first vertex which belongs to $r_i$ and $r_j$,
      we have $\lambda(v,r_i) - \lambda(v,r_j)= P$, where $P$ is an odd number smaller than $n$, the number of routes in ${\cal R_{\cal C}}$. In such a graph, there is no $(P,\tau)$-periodic affectation, because of $P < n$.\\
      If we consider a period of $2$, for all $i \neq j$, $\lambda(v,r_i) - \lambda(v,r_j) \mod 2 = 1$ . Therefore $(0,\dots,0)$ is a $(2,1)$-periodic affectation of ${\cal R}$.

@@

      In the context of cloud-RAN applications, we consider here the digraph $G=(V,A)$ modeling the target network 
      and two disjoint subsets of vertices $S$ and $T$ of equal cardinality, where $S$ is the set of RRHs and $T$ is the set of BBUs. 
      A \textbf{symmetric} assignment ${\cal C}$, is an involutive function from $S$ to $T$, which maps each element $s\in S$ to ${\cal C}(s)$ and ${\cal C}(s)$ to $s$. It can also be seen as the set of pairs $(s,{\cal C}(s))$ and $({\cal C}(s),s)$.\\
      The Routing function ${\cal R}$ associtates to each couple $(s,{\cal C}(s))$ a route called $r_s$ and to each couples $({\cal C}(s),s)$ a route called $r_{{\cal C}(s)}$.\\     
       We are given a period $P$, a routing function ${\cal R}$, and we consider a $(P,\tau)$-periodic affectation of ${\cal R}_{\cal C}$ which associates $m_s$ to the route $r_s$ which begins by $s$.  This affectation represents the following process: first a message is sent at $s$, through the route $r_s$, at time $m_s$.

@@

      \begin{center}
      \includegraphics[scale=0.3]{rrh.png}
      \end{center}

@@

       \begin{center}
      \includegraphics[scale=0.3]{BBU2.png}
      \end{center}

@@

  In this article, we works on given routages, but we could imagine a problem, in which, considering a given graph $G = (V,A)$ and a period P, the question is to find, if it exists, a routage graph ${\cal R}_{\cal C}$, in which there is $(P,\tau)$-periodic affectation.
\section{Solving PRA}
  \label{sec:complexity}
  \subsection{NP-Hardness}

 In this section we assume that the size of a message $\tau$ is equal to one. 
 We will prove the hardness of PRA and PALL for this parameter, which implies the hardness of the general problems. 
Consider an instance of the problem PRA, i.e., a routage graph $\cal{R_{\cal C}}$, a message size $\tau$ and a period $P$. \\
The {\bf conflict depth} of a route is the number of other routes which share an edge with it. 
The conflict depth of a routage graph  $\cal{R_{\cal C}}$ is the maximum of the conflict depth of the routes in $\cal{R_{\cal C}}$.\\
The {\bf load} of a routage graph is the maximal number of routes sharing the same arc.
It is clear that a $(P,\tau)$-periodic affectation must satisfy that $P$ is larger or equal to the load times $\tau$.

@@

      \Vertex[x=12,y=3]{t3}
      \Vertex[x=14,y=4]{t2}
      \Vertex[x=10,y=2]{t1}
      \tikzstyle{VertexStyle}=[shape = circle, draw, minimum size = 20pt]
	\tikzset{
      VertexStyle/.append style = {blue} }
	\Vertex[x=-8,y=3]{1}
	      \tikzset{
      VertexStyle/.append style = {green} }
	  \Vertex[x=-7,y=5]{2}

@@

      \SetVertexNoLabel
      \Vertex[x=2,y=5]{A}
      \Vertex[x=4,y=5]{B}
      \Vertex[x=10,y=5]{C}
      \Vertex[x=12,y=5]{D}
      \Vertex[x=6,y=3]{E}
      \Vertex[x=8,y=3]{F}
      \tikzset{
      EdgeStyle/.append style = {green} }
      \Edge(s2)(A)
      \Edge(A)(B)
      \Edge(B)(C)
      \Edge(C)(D)
      \Edge(D)(t2)

      \tikzset{
      EdgeStyle/.append style = {red} }
      \Edge(s3)(E)
      \Edge(E)(F)
      \Edge(F)(t3) 
	\tikzset{
      EdgeStyle/.append style = {blue} }
      \Edge(s1)(A)
      \Edge(A)(B)
      \Edge(B)(E)
      \Edge(E)(F)
      \Edge(F)(t1)

@@

      In this section, we consider graphs with a very simple topology that we call the {\bf star topology}. 
      First, for each arc $(u,v)$, there is also an arc $(v,u)$ with the same weight.
      Moreover, there is a special arc, the central arc, which is shared by all routes.
      All routes consist from an arc from its source to the central source node, denoted by {\bf $c_s$},
      then the central arc to {\bf $c_t$}, the central target node, and an arc to its target. In addition to those central nodes, there are two sets of vertices, $S$ and $T$, of cardinality $n$ and ${\cal C}$ a symmetric assignment from $S$ to $T$. 
      The routes are the directed paths of the form $s_i,c_s,c_t,t_i$ and $t_i,c_t,c_s,s_i$.

@@

      We assume the weight of the central arc to be $0$ since it appears in every route,
      its value does not matter when considering affectations. Collisions between messages can 
      only appear at node $c_s$ on the way forward and at node $c_t$ on the way back. \\
      Let us call {\bf forward period} and {\bf backward period} the time windows in respectively $c_s$ and $c_t$. A message issued by the source $s_1$ at time $m_{s_1}$ will reach $c_s$ at time $m_{s_1} + t(c_s,r_{s_1})$ in the forward period, and  $c_t$ at time $m_{{\cal C}(s_1)} + t(c_t,r_{{\cal C}(s_1)})$ in the backward period.
      A $(P,\tau)$-periodic affectation consist in choosing $\forall i \in n, m_i$, and $m_{{\cal C}(s_i)}$ such that, for any couples of routes $(r_{s_j}, r_{s_k}), j,k \in n$,
      we have $[t(c_s,r_j)] \cap [t(c_s,r_k)] = \emptyset$.
      and for any couples of routes $(r_{{\cal C}(s_j)}, r_{{\cal C}(s_k)}), j,k \in n$, we have $[t(c_s,r_{{\cal C}(s_j)})] \cap [t(c_s,r_{{\cal C}(s_k)})] = \emptyset$
%       
%   \todo{Définir la forward period et la backward period et dire à quoi correspond une affectation en terme de forward et backward period (placement de messages dans la première période}     
%       Therefore, 
%       finding an $(P,\tau)$-periodic affectation without collision, is equivalent to fix ...

@@

  In this subsection, we deal with a simpler version of the problem PALL.
  We ask for a $(P,\tau)$-periodic affectation with all waiting times equal to $0$. 
  In that case $T_{max}$ is not relevant anymore. Since $w_i=0$, choosing $m_i$, the offset of the route from
  $s_i$ to $t_i$, also sets the offset of the route from $t_i$ to $s_i$ to $m_i + \lambda(r_i)$.
  Moreover, in this context the weight of the arcs from $s_i$ to $c_s$ have no impact, therefore we assume they are 
  equal to $0$. Thus, in this case, we can consider only the forward and backward periods. The goal is now to put messages in the forward periods such that there is no collisions in the both periods.

@@

    \begin{proposition}
    There is an algorithm, which finds a $(P,\tau)$-periodic affectation in time $O(n^2)$ when $P \geq 3n\tau$.
    \end{proposition}
    \begin{proof}
     We consider the forward period and cut it into consecutive macro slots of time of size $\tau$. The algorithm works in the following way: select the first route for which the offset as not been chosen. Try every offset which put the message in a  
     free macro slot in the forward period. It also fixes the position of a message in the backward period, and we choose the first offset which does not create a collision in the backward period. 
     Assume we choose the offset of the route $r_{k+1}$, we have  at least $3n - k$ free macro-slots, since $P \geq 3n\tau$. Each of these $3n - k$ possible offset values translates into $3n - k$ positions of messages in the backward period. All these positions are separated by at least $\tau$ slots. There are already $k$ messages of size $\tau$ in the backward period. One such message can intersect at most $2k$ potential positions, therefore  amongst the possible $3n - k$ positions, there are  $3n - k -2k$ which are without collision. Since $k < n$, one of the choice is without collision, which proves that the algorithm terminates and find a
     $(P,\tau)$-periodic affectation. To ensure a complexity linear in $n$, it is enough to maintain a list of the free slices in the forward period and a list of the free intervals in the backward period. 
    \end{proof}

      \begin{center}
      \includegraphics[scale=0.3]{ex3nt.png}
      \end{center}

@@

	\begin{algorithm}[H]
	\caption{Greedy affectation}
	\begin{algorithmic}
	\REQUIRE ${\cal R}_{\cal C}$, period $P$
	\ENSURE A P-periodic affectation in p $\leq P$, or FAILURE
	\STATE $P1[P]$ slots of size $\tau$ in forward period.
	\STATE $P2[P]$ slots  of size $\tau$ in backward period.
	\FORALL{source i in S }

@@

	\end{algorithmic}
	\end{algorithm}
%       \paragraph{Period}
% 	This algorithm gives us a solution without waiting times in a maximum period $3.\tau.c$, if we have $c$ routes.\\
% 	Suppose that we have a period of $3.\tau.c$ slots, divided in $\tau$ macro-slots. Put a message in the first slot of size $\tau$ in the forward period, such that the corresponding area in the backward period is free.This message takes at most 2 slots of time $\tau$ in the backward period.\\

% 	When $k < c$ messages are put in the forward period, and we want to add another message, there is $3.c - k$ free slots of size $\tau$ in the forward period. Those $3.c - k$ gives us $3.c - k$ possible slots in the backward periods.\\
% 	The $k$ messages uses at most $2k$ slots of size $\tau$ used in the backward period. Since $k<c$,  $2k < 3.c - k$, thus using the pigeonhole principle, there is at least one free slot in the backward period for the new message.

    \subsubsection{Exhaustive generation}
	\begin{algorithm}[H]
	\caption{Exhaustive Generation}  
	\begin{algorithmic}
	\REQUIRE A routage graph ${\cal R}_{\cal C}$, period $P$, packet size $\tau$
	\ENSURE $(P,\tau)$-periodic affectation of ${\cal R}_{\cal C}$
	\STATE Forward-budget $\leftarrow$ $P$ - n * $\tau$
	\STATE Backward-budget $\leftarrow$ $P$ - n * $\tau$
	\STATE Free-Intervals $\leftarrow$ list of free intervals, init to $[0;P[$
	\FORALL{route i in  ${\cal R}_{\cal C}$}
	\FORALL{j in Free-Intervals }
	\IF{Message of the route i does not collides with scheduled routes}
	\STATE $m_i \leftarrow $ the first slot of Free-Intervals[j]
	\STATE Split the Free-Intervals considering the new packet
	\STATE Forward-budget $\leftarrow$ Forward-budget - {\em lost size}
	\STATE Backward-budget $\leftarrow$ Backward-budget - {\em lost size}
	\STATE call Exhaustive Generation on remaining routes
	\ENDIF
	\ENDFOR
	\ENDFOR
